Software Citation Principles

### 2016-07-23 sc-principles/rebuttal/rebuttal.tex

\niemnote{The idea of citations is not going to go away---but format and how they fit in ecosystem may change.}

\niemnote{Arfon: can you help here?}

\item \textbf{Reviewer comment:}
\emph{line 238: ``Citation is a record of software that is important to a research outcome'': This is another important issue, but not much more is said about this. What makes a software ``important'' to a research outcome? Or, more generally, what is the semantics of a software citation?}

\textbf{Our response:}
We have revised this discussion to better address these issues. In particular, we
added some references that discuss the importance of specific software versions to
research results (including data). For a particular study, it is up the authors to
determine which software is important, and peer-reviewers to decide if the authors have
done this sufficiently well, according to community standards, similarly to how citation
of papers in a paper is decided and judged.
\niemnote{how is this, including the new text?} \katznote{I've added a bit to the response, and lightly edited the text.  I also made a change in the text tied to reviewer 2 item 10 that I missed previously}

### 2016-07-23 sc-principles/software-citation-principles.tex
change in response to Kyle's work on reviewer 1 item 9

Some software which is, or could be, captured as part of data provenance may not be cited.
Citation is partly a record of software important to a research outcome\footnote{Citation
can be used for many purposes, including for software: which software has been used in
the work, which software has influenced the work, which software is the work superseding,
which software is the work disproving, etc.}, where provenance is a record of all steps
(including software) used to generated particular data within the research process.
Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
and thus may depend on the specific version
used~\cite{Sandve:2013gh,Wilson:2014aa}---furthermore, errors in software or
environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
This implies that for a data research product, provenance data will include some of
the cited software.
Similarly, the software metadata that is recorded as part of data provenance should be
a superset of the metadata recorded as part of software citation for the software that
was used in the work. The data recorded for reproducibility should also be a superset
of the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is
necessary for software citation for the purpose of software identification.
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }

>>>>

Some software which is, or could be, captured as part of data provenance may not be cited.
Citation is partly a record of software important to a research outcome\footnote{Citation
can be used for many purposes, including for software: which software has been used in
the work, which software has influenced the work, which software is the work superseding,
which software is the work disproving, etc.}, where provenance is a record of all steps
(including software) used to generated particular data within the research process.
Research results, including data, increasingly depend on software~\cite{Hannay:2009wp},
and thus may depend on the specific version
used~\cite{Sandve:2013gh,Wilson:2014aa}. Furthermore, errors in software or
environment variations can affect results~\cite{Morin:2012hz,Soergel:2015aa}.
This implies that for a data research product, provenance data will include some of
the cited software.
Similarly, the software metadata that is recorded as part of data provenance should be
will overlap the metadata recorded as part of software citation for the software that
was used in the work. The data recorded for reproducibility should also overlap
the metadata recorded as part of software citation.
These statements may also be true for software products.
In general, we intend the software citation principles to cover the minimum of what is
necessary for software citation for the purpose of software identification.
Some use cases related to citation (e.g., provenance, reproducibility) might have additional requirements beyond the basic metadata needed for citation, as Table~\ref{tab:use_cases} shows.
%\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
%\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
